Efficient and Effective Metric Space Indexing for Complete Record Linkage Al: Exploring approaches to probabilistic record linkage using similarity searching (metric space indexing?)

### 2017-11-11 paper.tex
Update on Overleaf.

% Peter 31 Oct: Current title does not really reflect the topic I
% think:
%\title{Exploring approaches to probablistic record linkage using
%        similarity searching}
% Peter 31 Oct: What about
\title{Efficient and Effective Metric Space Indexing for Complete
       Record Linkage
       Al: Exploring approaches to probabilistic record linkage using similarity searching (metric space indexing?) }
%  \thanks{The authors would like to thank the ...|

@@

something about record linkage and its importance, application examples

drawback of existing approaches: blocking and comparisons/classification
done separately, sometimes if bad linkage quality means to go back to redo
blocking.

@@

The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977. This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records. As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.

@@

The MIFile \cite{amato2014mi} approach is another inexact technique for performing similarity search. Using  MI_file,  each record is represented by the ordering of distances from a collection of reference objects. It is based on the idea that if two records are close to each other  will have similar neighbours and will thus be represented by similar of distances from the chosen reference objects. An inverted file is created mapping from...

>>>>

% Peter 31 Oct: Current title does not really reflect the topic I
% think:
%\title{Exploring approaches to probabilistic record linkage using
%        similarity searching}
% Peter 31 Oct: What about
\title{Efficient and Effective Metric Space Indexing for Complete
       Record Linkage
       Al: Exploring approaches to probabilistic record linkage using similarity searching (metric space indexing?) }
%  \thanks{The authors would like to thank the ...|

@@

% Peter, 20171111:

Record linkage, also known as entity resolution, data matching and
duplicate detection~\cite{Chr12}, is the process of identifying and
matching records that refer to the same real-world entities within and
across databases. The entities to be linked are often people (such as 
patients in hospital databases or customers in business databases),
but record linkage can also be applied to link consumer products or
bibliographic records~\cite{Chr12}. Commonly, record linkage is
challenged by the lack of unique entity identifiers (keys) in the
databases to be linked, which prevents the use of a database join.
Instead, the linkage of records requires the comparisons of the
attributes (or fields) available in common in the databases to be
linked. For databases that contain information about individuals,
the common attributes used often include the names, address,
dates of birth, and so on, of individuals.

To overcome data quality issues such as typographical errors and
variations (which are common in name and address values~\cite{Chr12}),
approximate string comparison functions (such as edit distance, the
Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used
to compare pairs of records, leading to a vector of similarities (one
similarity per attribute/field compared) for each pair. These
similarity vectors are then used to classify the compared record
pairs into matches (where it is assumed the records in a pair
correspond to the same real-world entity) and non-matches (where the
records are assumed to correspond to two different entities). Various
classification methods have been employed in record linkage, ranging
from simple threshold based to sophisticated clustering approaches
and supervised classifiers~\cite{}.

Besides the two challenges of a lack of unique entity identifiers and
data quality (which will affect linkage quality), record linkage is
challenged by the increasing sizes of the databases to be
linked~\cite{Don15} in today's Big Data era. To prevent the full
pair-wise comparison of all possible record pairs (which is quadratic
in the sizes of the databases to be linked), blocking, indexing and
filtering techniques are commonly used~\cite{Chr12}. The basic idea
of techniques is to split the databases into smaller blocks in a
computationally efficient way such that records that likely correspond
to the same entity are inserted into the same block. Only records that
are in the same block are then compared with each other in more
details. While blocking techniques facilitate the efficient linkage
of very large databases~\cite{Don15}, they generally have the drawback
of achieving scalability at the cost of (a possibly small) loss in
linkage quality, as potentially true matching record pairs are
removed in the blocking or indexing step. Blocking or indexing often
leads to a reduction in recall of the final linkage
result~\cite{Chr12}. A variety of blocking and indexing techniques,
as discussed in more details in the following section, have been
proposed, ranging from simple phonetic based blocking~\cite{Chr12} and
sorting of the databases~\cite{Dra12} to locality sensitive hashing
based techniques~\cite{Kim10} and even unsupervised~\cite{Kej13,Ram15}
or supervised~\cite{Bil06,Mic06} learning of optimal blocking schemes.

~ \\
\emph{Peter 20171111: Could we have 2 diagrams: traditional RL process
(block - compare - classify) versus our proposed 1 step approach
(metric indexing) to illustrate our idea?}
~ \\

The overall process of a record linkage application, where blocking
or indexing is preformed prior to the comparison and classification
steps as illustrated in Fig.~\ref{fig-rl-process}, adds a further
practical complexity. Often blocking, comparison and classification
are conducted using algorithms selected (and their parameters set)
based on technical and domain expertise of the record linkage user,
followed by a manual assessment of the linkage outcomes. If the
quality of the resulting links is not good enough for a certain
application, the linkage process needs to be repeated with different
parameter settings and possibly also alternative algorithms
selected. This can result in a time-consuming and labor intensive
process. A major challenge is often the heuristic nature of blocking,
where the choice of a certain blocking technique and its parameters
will determine the final outcome of a linkage. If a certain blocking
approach results in a loss of recall in the final linkage results,
the blocking needs to be redone, an often time consuming process.

In this paper we explore how existing blocking and indexing technique
can lead to loss in recall, even for record pairs that are highly
similar and likely correspond to true matches. As a result, using
such techniques can lead to \emph{incomplete} record linkage
results. To overcome this problem, we advocate the use of
\emph{metric space indexing} (MSI) approaches, where it is possible to
guarantee that all record pairs, up-to a certain similarity threshold,
are included in the record linkage results. Furthermore, MSI allows
us to combine the indexing and comparison steps (as traditionally
conducted independently) into one single step, making the overall
record linkage process both more efficient and more effective.

% peter 20171111: do we also need to mention we do not waste
% comparisons of record pairs with low similarities?

~ \\
\emph{Peter 20171111: here we need the figure showing that traditional
blocking techniques result in incomplete record linkage, while metric
indexing does not.}
~ \\

While MSI has been used extensively in areas such as database indexing
and general nearest neighbor search, its use in record linkage so far
has been limited. 
% Peter 20171111: a dangerous statement. if wrong we're stuffed.. 
% but I am not really aware of MSI approaches in RL..
Many different MSI approaches have been developed, such as various
metric tree structures~\cite{}. 
% Peter 20171111: needs same text from you guys on MSI..
In this paper, we will use the MSI techniques of M-tree and M-file (?)
and show how their use can facilitate an improvement of both
efficiency and effectiveness of the record linkage process.

\smallskip
\textbf{Contributions:} We make the following specific contributions
in the paper: (1) We investigate how existing indexing and blocking
techniques can result in \emph{incomplete} record linkage results,
as well as in an in-efficient record linkage process. (2) We explore
MSI for record linkage and propose a MSI based approach that achieves
\emph{complete}, efficient and effective record linkage. (3) We
evaluate our approach on several real-world data sets from diverse
domains and show its advantages over existing indexing and
blocking techniques for record linkage.

% old stuff below

@@

\iffalse  % *******************************************************
% we can't write this as it reveals us in the double-blind review
% we could ad to final paper
The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977. This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records. As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.
\fi  % *********************************************************

@@

The MIFile \cite{amato2014mi} approach is another inexact technique for performing similarity search. Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects. It is based on the idea that if two records are close to each other  will have similar neighbours and will thus be represented by similar of distances from the chosen reference objects. An inverted file is created mapping from...
